A Reimplementation of Python’s dict using Simple Tabulation Hashing and Linear Probing

### 2012-05-20 docs/paper.tex
update the theoretical results

\title{A Reimplementation of Python’s dict using Simple Tabulation Hashing and Linear Probing}
\author{
        Russell Cohen \\
            \and
        Thomas Georgiou \\
            \and
        Pedram Razavi
}
\date{\today}
\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{float}

@@

\subsection{Theoretical Results}
Recently, Patrascu et al. [2] showed that simple tabulation hashing provides unexpectedly strong guarantees and is constant competitive with a truly random hash function for hashing with linear probing. They also showed some experimental evaluations which proved that simple tabulation on fixed input length can be fast and robust when used with linear probing and cuckoo hashing.

@@

However, Patrascu et al. [2] describe simple tabulation hashing for fixed key widths.  We will need to figure out a variant of simple tabulation hashing for variable length strings that does not require a table for every possible index (we could potentially cycle through a fixed set of tables).  Performing table lookups is potentially an expensive operation with lots of cache hits mitigating the better cache behavior of linear probing so we will have to be careful about how we implement it.

@@

After implementation and profiling, we found that dictionary table lookups had
over twice as many level 1 cache misses with tabulation hashing than without.
Hypothesizing that the tabulation tables were evicting other, more relavant
data from the level 1 cache.  To fix this, we inserted x86 memory prefetch
instructions into our hash code to tell the CPU's caching system that it should
deprioritize storing it in cache.  We used GCC's
\begin{tt}\_\_builtin\_prefetch(void \*addr, char rw, char locality)\end{tt}
to do this.  However, this only yielded about a 1\% performance improvement.

@@

We attempted to measure the 
See Figure ~\ref{fig:tables}
 \begin{figure}[H]
   \centering
   \includegraphics[width=4in]{tables.pdf}
   \caption{}
   \label{fig:tables}
 \end{figure}

>>>>

\title{A Reimplementation of Python’s dict using Simple Tabulation Hashing and Linear Probing}
\author{
        Russell Cohen \\
            \and
        Thomas Georgiou \\
            \and
        Pedram Razavi
}
\date{\today}
\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{float}

@@

\subsection{Theoretical Results}
Recently, Patrascu et al. [2] showed that simple tabulation hashing provides unexpectedly strong guarantees and is constant competitive with a truly random hash function for hashing with linear probing. In their paper, they showed that for simple tabulation hashing with linear probing if the table size is $m = (1 + \epsilon)n$, then the expected time per operation is $O(1/\epsilon^2)$. This new guarantee asymptotically matches the
bound for a truly random function. They also showed some experimental evaluations which proved that simple tabulation on fixed input length can be fast and robust when used with linear probing and cuckoo hashing.

@@

However, Patrascu et al. [2] describe simple tabulation hashing for fixed key widths.  For implementing this new hashing scheme, first we needed to figure out a variant of simple tabulation hashing for variable length strings that does not require a table for every possible index (we could potentially cycle through a fixed set of tables).  Performing table lookups is potentially an expensive operation with lots of cache hits mitigating the better cache behavior of linear probing so we will have to be careful about how to implement this. In the following section we explain the implementation details of our approach to these problems.

@@

Cache

@@

We attempted to measure the 
 \begin{figure}[H]
   \centering
   \includegraphics[width=5in]{tables.pdf}
   \caption{}
 \end{figure}
